Enhancing the prediction of sales for newly published books by means of computational intelligence methods

### 2016-05-12 paper.tex
Sections 3 and 4 Improved

For this reason, usually the methods used to predict book sales have
been generally based on experts' experience.% Reference - JJ
 They analyse data about
sales and, taking into account their knowledge of industry, their
experience, and their perception about trends, could make decisions on
the companies production, i.e. how many books should be printed when a new
book is published. 
This is a complex problem, since the predictions are influenced by
external variables that must be taken into account, such as
seasonality, promotions, or fashions that expert managers might
subjectively apply \cite{Lapide1999,ChernWSF15}. % This
                                % has something to do with the
                                % previous paragraph? Put it there!!! 
%
  %[pedro] sobre la industria editorial no hay trabajos sobre predicción de ventas.
  % Lo habitual ha sido usar el conocimiento de expertos para realizar predicciones 
  %  o aconsejar acerca de la cantidad de libros a imprimir.
  % Lo más parecido que he encontrado son dos trabajos (uno en realidad) de Moon et al. 
  %  sobre "Use Blog Information As Book Sales Prediction":
  %     Moon2010ICSSSM
  %     Moon2010ICEC
Recently, some works have proposed analysing relationships between on-line information 
and book sales \cite{Moon2010ICSSSM,Moon2010ICEC}. Specifically, Moon
et al. propose using the number of blog references as an indicator of
the success of a book, and thus, its sales. Different sales patterns
are observed by analysing data obtained. 
% Antonio - describir esos 'sales patterns'. ;)
% FERGU: y también cuales son esos 'data obtained'
However, the authors used in these papers historical data from published books, 
so their proposed methodology can not be applied to the problem of unpublished books sales prediction.
%FERGU: pero nosotros también usamos historical data, no? qué diferencia hay con lo nuestro? por qué lo nuestro es lo mejor?
% (Paloma) Agree with Fergu, I don't think we should say this... Please review it.

@@

@@

In order to deal with this issue, i.e. forecasting sales of new products, some authors have used different data mining methods \cite{Hammond1990,Chang2009,ChingChin2010,Thomassey2012,Xia2012}. 
% Antonio - Maybe we should describe them briefly...
%
In the same line, Thomassey et al. \cite{SThomassey2014} proposed using a clustering method together with decision trees for sales forecasting in textile-apparel fashion industry, 
a very similar issue to the new book publishing one. % (Paloma) Very similar? Why?
The same problem was previously faced by Xia et al. \cite{Xia2012} using a 
forecasting method based on extreme machine learning with adaptive metrics of inputs.
% Antonio - of -> as???
% (Paloma) It's "of" in the title...
The complexity of this problem relies on the lack of historical data, on the short lifetime of the majority of items, 
% Antonio - ver si está bien el cambio 'large number' -> 'majority'
and on the influence of variables such as promotions, fashions, or economic environment \cite{Thomassey2012,Xia2012,SThomassey2014}, as mentioned above.
%FERGU: completar con: "In our case, these ideas ..."

@@

The main objective of this paper is to forecast sales for 
%FERGU: el objetivo es la herramienta o la predicción?
unpublished books for which no historical data are available. Instead, just some
%FERGU: pero usamos datos históricos de otros para la predicción, no? Como otros papers...
descriptive and variables are known in advance related to the new book. In
order to do that, we have used a training dataset which includes
historical sales data % (Paloma) so, have we or have we not historical data? It's all contradictory
 for a big amount of books published in Spain for
several years, with a set of 50 features or variables per pattern.

The initial dataset, provided in the framework of a contract by
Trevenque Editorial S.L., included 3159 books. Table
\ref{tabla:paramsOrig50} shows the initial set of variables describing
each book.

@@

% Antonio - briefly describe the variables

In this prediction problem there are other external variables that also affect 
the sales of a new book, such as those related to seasonality, trends, or 
promotions, that classical prediction methods can not deal with. %FERGU: por qué?
However, an expert can consider them to modify the forecast obtained by 
the automatic method. Even publishing companies tend to define rules or 
indexes to subjectively adjust the results of prediction based on
external data. 
Thus, the proposed method will analyse the provided data and
variables and will yield a forecast of sales according to them. This
result could be used as a reference for a human expert, who would
consider it in his or her own prediction process.

% This section is too short and really does not describe the problem
% to its full extent. First, it should say if some of these variables
% were discarded for some reason. Group the variables we have in those
% available to publisher and those that can be used afterwards. And
% then give an idea of the complexity of the problem and how difficult
% it should be to find a solution - JJ

There is not a standard methodology to solve the general problem of sales forecasting. This is why the steps we have followed in this paper are those usually applied in a Knowledge Discovery in Databases (KDD) process \cite{fayyad1996data}. These are described below.
% Antonio - quito esto porque ya no es lo que se ha hecho. El objetivo del paper no es la metodología a seguir, eso es estándar. ;)
%FERGU: hacer más hincapié de esto en la intro entonces
%In this work we have followed these steps: %FERGU: por qué estos steps y no otros? Ponerlo aquí. Esta metodología es general o solo para este paper?
%\hspace{2cm}
%\begin{verbatim}
%  Step 1: Collect and analyse data
%       - Data preprocessing
%       - Feature selection
%  Step 2: Select the best forecasting method
%       - Evaluate forecasting models
%       - Perform forecast
%  Step 3: Subjectively adjust obtained results
%      - Apply additional expert knowledge, such as information 
%        about seasonality or trends.
%\end{verbatim}

%Usually, a preliminary analysis of the data is carried out to eliminate those 
%variables that describe the product but do not provide useful information during 
%the prediction process.

First, after the publishing company has provided the raw data
(extracted directly from its database), a preprocessing method has
% This should go to the previous section. 
been performed in order to remove incorrect or incoherent patterns,
such as those with null values in required variables, or with negative
sales, for instance. % (Paloma) Why remove those with null values and not to use an algorithm to fill those missing values?
Furthermore, the initial set of variables, presented
in Table \ref{tabla:paramsOrig50}, has been reduced by an expert to
use just those that can be known in advance when a new book is
launched, i.e. \textit{pre-sales data}. % (Paloma) Where are they? In table 2?

% You have to present the problem and not the solution. It's not "what
% you did last summer", it is SCIENCE. Why did you think FS was
% needed? What procedure did you use for it?
Then, a Feature Selection (FS) process \cite{kittler1986feature} has
been carried out, as it is an effective dimensionality reduction
technique and an essential preprocessing method to discard `noisy'
(non-useful) features \cite{Krishnapuram2004}. The aim of these
% Did you know there were noisy features? Why? 
% You should include in the previous section an exploratory data
% analysis that describes the range of variables, how many categories
% those variables have, and so an do forth - JJ 
algorithms is exploring the space of attributes to find out which
subset of them yields the best results in a classification or
prediction problem. 
  In this stage, both numerical and visualisation-based approaches have been applied, in order to select the best variables for the forecasting 
methods.  
% Antonio - Decidir si es 'forecasting' o 'regression' (o si son indiferentes) y usarlo siempre en el texto
  This task may also help to identify leading indicators to be considered by a human expert to make predictions.

After this process, several forecasting techniques have been tested, such as rule- or tree-based, together with numerical methods. 
% Antonio - is 'numerical' the best term for these kind of methods?
%FERGU: OJO que en esta sección a veces usáis el presente perfecto (have been) y otras el presente (is performed)
% Antonio - Corregido. ;)
%

Once the models and results have been obtained for the whole dataset, the same methodology has been applied to some specific cases of publishing companies in order to prove the value of the proposal.
Some specific interpretable models have been created as a final result of the work. The aim is that these can be used as a decision-aid tool for the experts who have to decide on the best print-run quantity for new books to be published, in order to either maximise the benefit, or minimise the costs.

%The created forecasting models and results, along with the selected subset of features, are analysed and presented to the experts in order to aid his/her decision. They might manually adjust the forecasting results taking into account their knowledge in industry experience and their perception about trends.

%FERGU: poner aquí algo de un párrafo que explique como se valida la metodología
% Antonio - TODO: esto habrá que explicarlo mejor cuando esté hecho y veamos lo que sale. ;)

%-------------------------------------------------
\subsection{Data collection}
\label{subsec:data_collection}

Due to this situation, the initial set of features in the dataset must be revised in order to consider just those that can be known before the launch of the book, i.e. those whose value do not depend on the book sales. Therefore, we have selected the subset of features described in Table \ref{tabla:params_pre_sales}.

\begin{table}
% Shouldn't this table go to "Problem description"? It is a
% preliminary selection - JJ
\caption{Considered pre-sales features and their types. The third column shows
  the names of the variables used in the applied methods, and the last one indicates if they have been used as an input (independent) or an output (dependent) variable.} 
\label{tabla:params_pre_sales}
\begin{center}
\begin{tabular}{|c|l|l|c|c|}
\hline 
No. & Feature Name & Variable & Type & In/Out\\
\hline 
1 & retail price (when launched) & \texttt{ret\_price} & numerical & input\\
2 & main subject (code) & \texttt{subject1} & numerical & input\\
3 &  publisher + collection (joint code) & \texttt{collection} & categorical & input\\
4 & bookbinding (code) & \texttt{binding} & categorical & input\\
5 & gifts (promotional books) & \texttt{gifts} & numerical & input\\
6 & units distributed as novelty & \texttt{distrib\_novelty} & numerical & input\\
7 & total number of points of sale & \texttt{tot\_points\_sale} & numerical & input\\
8 & total number of points of sale (1st year) & \texttt{tot\_points\_sale\_1st\_year} & numerical & input\\
9 & weeks on sale & \texttt{weeks\_sale} & numerical & input\\
10 & print run & \texttt{print\_run} & numerical & input\\
\hline 
\hline
11 & total sales & \texttt{total\_sales} & numerical & output\\
\hline 
\end{tabular}
\end{center}
\end{table}

@@

With regard to the feature \textit{publisher}, we have removed it from the final dataset since it is indeed included in the code of \texttt{collection}. In addition, we think it is not really relevant for the accuracy of the models and, moreover, publishers will not consider the results in other publishing companies for predicting its own sales.
% Antonio - Think about this... should we mention this?
% (Paloma) If you finally leave it, please review it because this last sentence is a bit confusing

@@

%-------------------------------------------------
\subsection{Data preprocessing and initial analysis}
\label{subsec:data_preproces}

@@

At this step, the developed forecasting models are evaluated and assessed based on different evaluation criteria. In addition to the \textit{correlation coefficient}, we have considered the widely accepted standard formulae in the literature \cite{gepsoft} (Equations \ref{eq:MAE} to \ref{eq:RRSE}), which are implemented in tools like Weka \cite{pentaho,Witten2011} or R \cite{otexts,Hyndman2013}.
Specifically, in order to compute the forecasting errors of the different methods, Mean absolute error (MAE), Root mean squared error (RMSE), Relative absolute error (RAE) and Root relative squared error (RRSE) have been selected:
%FERGU: have been selected for this methodology, for this work... algo así
% Antonio - ya se ha dicho. ;)

@@

%La investigación descrita en este trabajo se ha llevado a cabo conjuntamente con una empresa editorial que ha facilitado un conjunto de datos con los que trabajar, así como con la ayuda de un experto en ventas de este campo para validar los resultados obtenidos.
A dataset consisting of 3159 books provided by the Spanish %FERGU: mover este párrafo debajo del siguiente
publishing company Trevenque Editorial S.L. has been used.

>>>>

For this reason, usually the methods used to predict book sales have
been generally based on experts' experience.% Reference - JJ
 They analyse data about sales and, taking into account their knowledge of industry, their experience, and their perception about trends, could make decisions on the companies production, i.e. how many books should be printed when a new book is published. 
This is a complex problem, since the predictions are influenced by
external variables that must be taken into account, such as
seasonality, promotions, or fashions that expert managers might
subjectively apply \cite{Lapide1999,ChernWSF15}. % This
                                % has something to do with the
                                % previous paragraph? Put it there!!! 
%
  %[pedro] sobre la industria editorial no hay trabajos sobre predicción de ventas.
  % Lo habitual ha sido usar el conocimiento de expertos para realizar predicciones 
  %  o aconsejar acerca de la cantidad de libros a imprimir.
  % Lo más parecido que he encontrado son dos trabajos (uno en realidad) de Moon et al. 
  %  sobre "Use Blog Information As Book Sales Prediction":
  %     Moon2010ICSSSM
  %     Moon2010ICEC
Recently, some works have proposed analysing relationships between on-line information and book sales \cite{Moon2010ICSSSM,Moon2010ICEC}. Specifically, Moon et al. propose using the number of blog references as an indicator of
the success of a book, and thus, its sales. Different sales patterns
are observed by analysing the obtained data. 
% Antonio - describir esos 'sales patterns'. ;)
% FERGU: y también cuales son esos 'data obtained'
%The authors considered in these papers historical data from published books, 
%so their proposed methodology can not be applied to the problem of unpublished %books sales prediction.
%FERGU: pero nosotros también usamos historical data, no? qué diferencia hay con lo nuestro? por qué lo nuestro es lo mejor?
% (Paloma) Agree with Fergu, I don't think we should say this... Please review it.

@@

% Antonio - He intentado escribirlo, pero no me queda claro... Los datos que tenemos son también de libros ya publicados, ¿no?. Estamos considerando variables pre-venta sólamente, eso sí, pero usando datos históricos queremos predecir ventas futuras.
% Quizá lo que habría que señalar (aquí y en los demás sitios que se dice lo de los datos históricos) es lo de las variables que se consideran, si son variables post-venta o pre-venta, como es nuestro caso. ;)

@@

In order to deal with this issue, i.e. forecasting sales of new products, some authors have used different data mining methods \cite{Hammond1990,Chang2009,ChingChin2010,Thomassey2012,Xia2012}. 
% Antonio - Maybe we should describe them briefly...
%
In the same line, Thomassey et al. \cite{SThomassey2014} proposed using a clustering method together with decision trees for sales forecasting in textile-apparel fashion industry, 
a very similar issue to the new book publishing one. % (Paloma) Very similar? Why?
The same problem was previously faced by Xia et al. \cite{Xia2012} using a 
forecasting method based on extreme machine learning with adaptive metrics of inputs.

The complexity of this problem relies on the lack of historical data, on the short lifetime of the majority of items, 
and on the influence of variables such as promotions, fashions, or economic environment \cite{Thomassey2012,Xia2012,SThomassey2014}, as mentioned above.
%FERGU: completar con: "In our case, these ideas ..."

@@

The problem to solve in this work is the forecasting of sales for new launched books in the Spanish market, considering a set of `pre-sales variables'.
The objective is to offer a tool for publishers which will aid them to decide the best print run for a new book which before its publication process starts. This is a forecasting problem for which no specific historical data are available, since the same book has not been previously published.
However, we have considered data about other previously published books and have taken as reference a set of variables (or features), related with the book publication process.

The training dataset included historical data about 6083 books published in Spain for several years, with an initial set of 50 variables per pattern. It was provided by the Spanish publishing company Trevenque Editorial S.L.. Table
\ref{tabla:paramsOrig50} shows the initial set of variables describing
each book.

@@

Starting from these raw data, a preprocessing method has been performed in order to remove incorrect or incoherent patterns, such as those with null values in required variables, or with negative sales, for instance. 
% (Paloma) Why remove those with null values and not to use an algorithm to fill those missing values?
% Antonio - because we should not fill important variables automatically, this adds noise to the data, so I think it is better to remove these patterns, since there are still a lot in the dataset. 
% Indeed, there were just a couple of patterns without important values, that probably were errors of the extraction process or bad insertions in the database.
Furthermore, the initial set of variables, presented in Table \ref{tabla:paramsOrig50}, has been reduced by an expert to use just those that can be known before a new book is launched, i.e. \textit{pre-sales data}. Thus, their value do not depend on the book sales. The selected the subset of features is described in Table \ref{tabla:params_pre_sales}.

\begin{table}
% Shouldn't this table go to "Problem description"? It is a
% preliminary selection - JJ
% Antonio - done
\caption{Considered pre-sales features and their types. The third column shows
  the names of the variables used in the applied methods, and the last one indicates if they have been used as an input (independent) or an output (dependent) variable.} 
\label{tabla:params_pre_sales}
\begin{center}
\resizebox{12cm}{!}{
\begin{tabular}{|c|l|l|c|c|}
\hline 
No. & Feature Name & Variable & Type & In/Out\\
\hline 
1 & retail price (when launched) & \texttt{ret\_price} & numerical & input\\
2 & main subject (code) & \texttt{subject1} & numerical & input\\
3 &  publisher + collection (joint code) & \texttt{collection} & categorical & input\\
4 & bookbinding (code) & \texttt{binding} & categorical & input\\
5 & gifts (promotional books) & \texttt{gifts} & numerical & input\\
6 & units distributed as novelty & \texttt{distrib\_novelty} & numerical & input\\
7 & total number of points of sale & \texttt{tot\_points\_sale} & numerical & input\\
8 & total number of points of sale (1st year) & \texttt{tot\_points\_sale\_1st\_year} & numerical & input\\
9 & weeks on sale & \texttt{weeks\_sale} & numerical & input\\
10 & print run & \texttt{print\_run} & numerical & input\\
\hline 
\hline
11 & total sales & \texttt{total\_sales} & numerical & output\\
\hline 
\end{tabular}
}
\end{center}
\end{table}

@@

With regard to the feature \textit{publisher}, it has been removed from the final dataset since it is indeed included in the code of \texttt{collection}. In addition, we think it is not really relevant for the accuracy of the models and, moreover, publishers probably will not want to use the results of other publishing companies for predicting their own sales.
% Antonio - Think about this... should we mention this?
% (Paloma) If you finally leave it, please review it because this last sentence is a bit confusing
% Antonio - I guess it is now clearer

@@

In this prediction problem there are other external variables that also affect 
the sales of a new book, such as those related to seasonality, trends, or 
promotions, that classical prediction methods can not deal with, since they are not included in the dataset due to the difficulty to model them. %FERGU: why?
% Antonio - explained.
However, an expert can consider them to modify the forecast obtained by 
the automatic method. Even publishing companies tend to define rules or 
indexes to subjectively adjust the results of prediction based on
external data. 
Thus, the proposed method will analyse the provided data and
variables and will yield a forecast of sales according to them. This
result could be used as a reference for a human expert, who would
consider it in his or her own prediction process.

% This section is too short and really does not describe the problem
% to its full extent. First, it should say if some of these variables
% were discarded for some reason. Group the variables we have in those
% available to publisher and those that can be used afterwards. And
% then give an idea of the complexity of the problem and how difficult
% it should be to find a solution - JJ
% Antonio - I have included much more contents form other sections. could you do the rest?

% You should include data visualization in problem description,
% because it shows the distribution of problem variables and then
% suggests ways to deal with it - JJ
% You should include in the section an exploratory data
% analysis that describes the range of variables, how many categories
% those variables have, and so an do forth - JJ
% Antonio - as you know SOM has been applied later as a feature selection method... I don't know if we could reference the figure of SOM here or the other way round (do the analysis here and reference later), but in that case, we should explain SOM before, which could be a mess...
% Antonio - If you meant using other visualization methods... could you do it also?

There is not a standard methodology to solve the general problem of sales forecasting. This is why the steps we have followed in this paper are those usually applied in a Knowledge Discovery in Databases (KDD) process \cite{fayyad1996data}. These are described in the following sections.

As a summary, after the data has been collected and initially preprocessed (as described in previous section), other preprocessing techniques are applied, such as Feature Selection (FS) \cite{kittler1986feature} (see Section \ref{subsec:data_preproces}). These methods check the relevance of all the variables of the patterns, aiming to reduce even more this set of features, in order to obtain easier-to-interpret models, in a less amount of time (due to their lower complexity), and with a fair or, hopefully, better accuracy.
This task may also help to identify leading indicators to be considered by a human expert to make predictions.

% You have to present the problem and not the solution. It's not "what
% you did last summer", it is SCIENCE. Why did you think FS was
% needed? What procedure did you use for it?
% Antonio - done
In this stage, both numerical and visualisation-based approaches have been applied, in order to select the best variables for the forecasting 
methods.  
% Antonio - Should we speak about 'forecasting' or 'regression', Sam? Coudl they be used with the same meaning?

After this process (in Section \ref{subsec:select_method}), a selection of the best forecasting techniques have been conducted, looking for those rule- or tree-based, together with numerical methods which could yield the best results.

Once the models and results have been obtained for the whole dataset (Section \ref{sec:experiments}), the same methodology has been applied to some specific cases of publishing companies in order to prove the value of the proposal (Section \ref{sec:example_cases}).
Some specific interpretable models have been created as a final result of the work. The aim is that these can be used as a decision-aid tool for the experts who have to decide on the best print-run quantity for new books to be published, in order to either maximise the benefit, or minimise the costs.

%-------------------------------------------------
%\subsection{Data collection}
%\label{subsec:data_collection}

%-------------------------------------------------
\subsection{Data preprocessing by feature selection}
\label{subsec:data_preproces}

Then, a Feature Selection (FS) process  has
been carried out, as  It is an effective dimensionality reduction
technique and an essential preprocessing method to discard `noisy'
(non-useful) features \cite{Krishnapuram2004}.

The aim of these
% Did you know there were noisy features? Why?

algorithms is exploring the space of attributes to find out which
subset of them yields the best results in a classification or
prediction problem.

@@

At this step, the developed forecasting models are evaluated and assessed based on different evaluation criteria. In addition to the \textit{correlation coefficient}, we have considered the widely accepted standard formulae in the literature \cite{gepsoft} (Equations \ref{eq:MAE} to \ref{eq:RRSE}), which are implemented in tools like Weka \cite{pentaho,Witten2011} or R \cite{otexts,Hyndman2013}.
Specifically, in order to compute the forecasting errors of the different methods, Mean absolute error (MAE), Root mean squared error (RMSE), Relative absolute error (RAE) and Root relative squared error (RRSE) have been selected:

@@

%La investigación descrita en este trabajo se ha llevado a cabo conjuntamente con una empresa editorial que ha facilitado un conjunto de datos con los que trabajar, así como con la ayuda de un experto en ventas de este campo para validar los resultados obtenidos.
A dataset consisting of 6000 books provided by the Spanish %FERGU: mover este párrafo debajo del siguiente
publishing company Trevenque Editorial S.L. has been used.
